A quantum parallel Markov chain Monte Carlo

### 2022-08-01 main.tex
just need citation

\begin{figure}[!t]
	\centering
	\scalebox{0.9}{
		\begin{minipage}{1\textwidth}
			\begin{algorithm}[H]
	\caption{Quantum exponential searching algorithm \citep{boyer1998tight}}\label{alg:expoSearch}
	\KwData{An oracle gate $\U_f$ taking $\ket{x}\ket{y}$ to $\ket{x}\ket{y\oplus f(x)}$ for a function $f(x):\{0,\dots,N-1\}\rightarrow \{0,1\}$ with unknown number of solutions; $n=\log_2(N)$.}
	\KwResult{If a solution exists, an $n$-bit binary string $x_0$ satisfying $f(x_0)=1$; if no solution exists, the algorithm runs forever.}
	$m \gets 1$\;
	$\gamma \gets 6/5$\;
	success $\gets$ \textcolor{blue}{FALSE}\;
	\While{\emph{success}$\neq$\emph{\textcolor{blue}{TRUE}}}{
		$j \gets Uniform \{0,\cdots,m-1\}$\;
		$\ket{0}^{\otimes n}\ket{1} \longrightarrow \H^{\otimes n+1}\ket{0}^{\otimes n}\ket{1}= \ket{h}\ket{-}$\; 
		$\ket{h}\ket{-} \longrightarrow \G^j \ket{h}\ket{-} = \ket{x}\ket{-}$;\hspace{1em}  \Comment{$j$ Grover iterations from Alg \ref{alg:grover}.}  
		$\ket{x} \longrightarrow x$; \hspace{16em} \Comment{Measure and check.}  
		\uIf{$f(x)=1$}{$x_0 \gets x$\;
			success $\gets$  \textcolor{blue}{TRUE}\;}
		\Else{$m \gets \min\left( \gamma m, \sqrt{N}\right)$;\hspace{4em}  \Comment{Increase $m$ in case of failure.}
	}}
	\Return{$x_0$}\ .

@@

 \begin{figure}[!t]
 	\centering
 	\includegraphics[width=1\linewidth]{searchingFigs.pdf}
 	\caption{[Left] Total number of oracle evaluations required by quantum exponential searching algorithm (Algorithm \ref{alg:expoSearch}) for different numbers of solutions $M$ and search set sizes $N$ from 500 independent simulations each.  Horizontal lines at $\frac{9}{4}\sqrt{N}$ represent upper bounds on expected total number of evaluations to obtain a solution for the $M=1$ problem. [Right] Probabilities of success for the fixed-point quantum search algorithm \citep{yoder2014fixed} for different proportions of solutions $\lambda=M/N$ and selecting different lower bounds $w$ on $M/N$ with error tolerance $\delta^2$. }\label{fig:expo}
 \end{figure}

@@

We have shown how parallel MCMC may enjoy quadratic speedups by combining quantum minimization with the Gumbel-max trick.  Within a QPMCMC iteration, the current state represents a warm-start for the sampling-turned-optimization problem, leading to increased efficiencies for the quantum minimization algorithm.  This work leads to two immediate questions.  First, more recent quantum minimization algorithms exist \citep{baritompa2005grover,farhi2014quantum}.  Would these confer additional speedups for quantum parallel MCMC and benefit from warm-starting in a similar manner? Second, there are other models and algorithms in statistics, machine learning and statistical mechanics that require sampling from potentially costly discrete distributions.  Can one adapt  the proposed strategies to these?  Large-scale, practical quantum computing is still a long way off, but quantum algorithms are ripe for mainstream computational statistics.

>>>>

\begin{figure}[!tp]
	\centering
	\scalebox{0.9}{
		\begin{minipage}{1\textwidth}
			\begin{algorithm}[H]
	\caption{Quantum exponential searching algorithm \citep{boyer1998tight}}\label{alg:expoSearch}
	\KwData{An oracle gate $\U_f$ taking $\ket{x}\ket{y}$ to $\ket{x}\ket{y\oplus f(x)}$ for a function $f(x):\{0,\dots,N-1\}\rightarrow \{0,1\}$ with unknown number of solutions; $n=\log_2(N)$.}
	\KwResult{If a solution exists, an $n$-bit binary string $x_0$ satisfying $f(x_0)=1$; if no solution exists, the algorithm runs forever.}
	$m \gets 1$\;
	$\gamma \gets 6/5$\;
	success $\gets$ \textcolor{blue}{FALSE}\;
	\While{\emph{success}$\neq$\emph{\textcolor{blue}{TRUE}}}{
		$j \gets Uniform \{0,\cdots,m-1\}$\;
		$\ket{0}^{\otimes n}\ket{1} \longrightarrow \H^{\otimes n+1}\ket{0}^{\otimes n}\ket{1}= \ket{h}\ket{-}$\; 
		$\ket{h}\ket{-} \longrightarrow \G^j \ket{h}\ket{-} = \ket{x}\ket{-}$;\hspace{1em}  \Comment{$j$ Grover iterations from Alg \ref{alg:grover}.}  
		$\ket{x} \longrightarrow x$; \hspace{16em} \Comment{Measure and check.}  
		\uIf{$f(x)=1$}{$x_0 \gets x$\;
			success $\gets$  \textcolor{blue}{TRUE}\;}
		\Else{$m \gets \min\left( \gamma m, \sqrt{N}\right)$;\hspace{4em}  \Comment{Increase $m$ in case of failure.}
	}}
	\Return{$x_0$}\ .

@@

 \begin{figure}[!tp]
 	\centering
 	\includegraphics[width=1\linewidth]{searchingFigs.pdf}
 	\caption{[Left] Total number of oracle evaluations required by quantum exponential searching algorithm (Algorithm \ref{alg:expoSearch}) for different numbers of solutions $M$ and search set sizes $N$ from 500 independent simulations each.  Horizontal lines at $\frac{9}{4}\sqrt{N}$ represent upper bounds on expected total number of evaluations to obtain a solution for the $M=1$ problem. [Right] Probabilities of success for the fixed-point quantum search algorithm \citep{yoder2014fixed} for different proportions of solutions $\lambda=M/N$ and selecting different lower bounds $w$ on $M/N$ with error tolerance $\delta^2$. }\label{fig:expo}
 \end{figure}

@@

We have shown that parallel MCMC enjoys quadratic speedups by combining quantum minimization with the Gumbel-max trick.  Within a QPMCMC iteration, the current state represents a warm-start for the sampling-turned-optimization problem, leading to increased efficiencies for the quantum minimization algorithm.   Moreover, combining this approach with the Tjelmeland correction \citep{holbrook2021generating} results in a total complexity reduction from $\mathcal{O}(P^2)$ to $\mathcal{O}(\sqrt{P})$.  Preliminary evidence suggests that our strategy may find use when target distributions exhibit extreme multimodality.

This work leads to two interesting questions.   First, what is the status of inference obtained by QPMCMC?  The QPMCMC selection step relies on quantum minimization, an algorithm that only achieves success with probability $1-\epsilon$.  While empirical studies suggest that this error induces little bias, it would be helpful to use this $\epsilon$ to bound the distance between the target distribution and the distribution generated by QPMCMC.  Such theoretical efforts would need to extend recent formalizations of the parallel MCMC paradigm \textcolor{red}{CITATION} to account for biased MCMC kernels.

Second, there are other models and algorithms in statistics, machine learning and statistical mechanics that require sampling from potentially costly discrete distributions.   Can one adapt  our quantum Gumbel-max trick to these?  Approximate Bayesian computation \citep{csillery2010approximate} extends probabilistic inference to contexts within which one only has access to a complex, perhaps computationally intensive, forward model.  Having sampled model parameters from the prior, one accepts or rejects these parameter values based on the discrepancy between the observed and simulated data.  If one succeeds in embedding forward model dynamics within a quantum circuit, then one may plausibly select from many parameter values using our quantum Gumbel-max trick.

Large-scale, practical quantum computing is still a long way off, but quantum algorithms are ripe for mainstream computational statistics.

%This work leads to two immediate questions.  First, more recent quantum minimization algorithms exist \citep{baritompa2005grover,farhi2014quantum}.  Would these confer additional speedups for quantum parallel MCMC and benefit from warm-starting in a similar manner? Second, there are other models and algorithms in statistics, machine learning and statistical mechanics that require sampling from potentially costly discrete distributions.  Can one adapt  the proposed strategies to these?  Large-scale, practical quantum computing is still a long way off, but quantum algorithms are ripe for mainstream computational statistics.
